### 2016-02-08 paper.tex
Minor changes in Sec 3

\begin{abstract}
As electronic devices get increasingly personalized,
mixed signal integrated circuits such as analog to digital converters (ADCs)
and digital to analog converters (DACs) play a pivotal role in interfacing 
between the real world and the digital domain. Side-by-side, the security
requirements of these devices are increasing manifold. Many of these devices
require to be authenticated and have built in techniques to support 
encryption and prevent counterfeiting. For embedded devices, 
these requirements are efficiently implemented by 
using physically unclonable functions (or PUFs) that make use of the 
unique nanoscale variations during device fabrication.

In this paper, we show how mixed signal ICs in a device like ADCs and DACs 
have properties that make them amicable for use as intrinsic PUFs. By
virtue of having analog components, the PUF properties in these ICs
are more stronger compared to digital PUFs such as those based on
SRAM and ring oscillators. An added advantage is that mixed signal PUFs
have considerably less overheads compared to the contemporary digital PUFs. 
\end{abstract}

\section{Introduction}
The physical characteristics of every electronic device is unique.
The uniqueness is due to micro and nanoscale variations in the 
silicon substrate and fabrication process. A physically unclonable function 
(or PUF) uses this uniqueness to provide a cryptographic key that 
cannot be forged. PUFs have wide application, especially for device
authentication, anti-counterfeiting, and tamper resistance.

Several PUFs have been constructed over the years. Most of them
from electronic components such as the SRAM, ring oscillators, flip flops,
FPGA units etc. A few non-electronic PUFs such as optical, acoustic,
and magnetic PUFs have also been proposed. The important requirements
for a PUF is that they should be compact, unforgeable, unpredictable, 
irreversible, reliable, and easy to evaluate. These requirements are 
easier said than met. Many of the requirements are contrasting, while 
many are drastically affected by noise and other environmental factors 
such as the ambient temperature. Thus the search for the search for 
the perfect PUF is an open problem and active area of research.

We propose to construct PUFs using mixed-signal components such as {\em analog to
digital converters} (ADCs) and {\em digital to analog converters} (DACs).
As their names suggest, ADCs convert analog voltage signals 
to equivalent digital data, while DACs convert digital data into the corresponding
analog voltage. These devices are omnipresent in embedded 
systems such as mobile phones, wireless sensor nodes, smart tablets, etc.
and are often integrated into micro-controllers~\cite{xxx}. Several 
micro-controller vendors have multiple DACs and ADCs built into a
single device.

Both ADCs and DACs have analog as well as digital parts. This is
appealing for constructing PUFs due to the following reasons.
\begin{itemize}
\item As ADCs and DACs have digital parts that are interfaced with the
processor, the resulting PUFs built from them are intrensic and  
directly accessible by applications running in the system. No external 
measuring equipment is required.

\item By virtue of having analog components in mixed-signal devices, 
micro and nano scale variations are likely to be observed in two dimensions : 
device delays and voltage levels. Compare this with the contemporary intrensic
PUFs, where the variations are only in one dimension; either in the delays or
the voltage level. For instance, in delay based PUFs like the arbiter and the 
ring oscillator PUF, inherient differences
only in delays are captured. On the other hand in memory based PUFs, memories 
cells settling to either 0 or 1 is captured. In comparison, there are multiple 
values an analog voltage can take.

\item MS-PUFs do not require any dedicated circuits on chip. It reuses ADCs and
DACs that are present in the microprocessor. Thus MS-PUFs do not have any overheads.
\end{itemize}

\begin{figure}[!b]
\centering
\caption{Mixed-Signal PUF in a Micro-controller with built-in ADC and DAC.\label{fig:mspuf} }
\includegraphics[width=6cm]{figs/mspuf}
\end{figure}

The scheme for the MS-PUF is shown in Figure~\ref{fig:mspuf}. Outputs of DACs in the device
are fedback to the inputs of ADCs to form feedback channels. 
The DAC in each channel would receive digital data from the processor and convert it
to an equivalent analog value. The ADC at the other end of the channel would reconvert
the analog value to a corresponding digital data. The reconverted digital data would 
be affected by several factors including quantification and environmental noise,
and inherient variations due to the fabirication of the ADCs and DACs. The latter
aspect is useful for constructing PUFs.

The PUF {\em challenge} is a 
digital time-series pattern and a random sequence of feedback channel identifiers. 
The time-series pattern is transferred to the first channel in the sequence and the 
and the result read from the corresponding ADC. This result is a time-series 
digital pattern and is sent to the next channel in the sequence. This `looping' of the
pattern continues for every channel in the sequence. The result after the final loop
is taken as the PUFs {\em response}.

The paper \ldots

The structre of the paper is as follows....

%\bibliographystyle{abbrv}
%\bibliography{screfs}

\section{Background}
\subsection{Physically Unclonable Functions}
A PUF relies on the fact that two similar devices tend to exhibit
different behaviour because of differences in their physical microstructure
and process variations during manufacture. A PUF extracts these 
differences by a 
series of challenges and responses. Each response is a 
function of the challenge and the device characteristics. Thus each device will
have a different response for the same challenge. A collection of challenge-response pairs
are used for several cryptographic applications such as device fingerprinting, 
authentication, anti-cloning, and generation of cryptographic keys.

The PUF's challenge and response can be represented by a function $f$ 
mapping the domain of challenges ($\mathcal C$) to a range of responses ($\mathcal R$).
\begin{equation*}
f : \mathcal C \rightarrow \mathcal R \,\,\,
  | \,\,\, f(c) = r, \text{ where } c \in \mathcal C \text{ and } r \in \mathcal R
\end{equation*}

The function $f$ should be one-way, easy to evaluate and reproduce, yet unpredictable
and unclonable.

\subsection{Mixed Signal Circuits}
Mixed signal circuits interface the analog and digital domains in a chip. 
They are typically used to digitalize real world signals captured by
a sensor, which
can then be processed by software. They are also used to actuate external 
hardware devices. 
For instance, in a mobile phone, audio signals captured from 
a microphone are digitalized by means of an {\em analog to digital converter
(ADC)}, before being processed by the mobile software. Received audio data is
converted to analog by a {\em digital to analog converter (DAC)} and actuate the
speakers of the mobile phone. This section provides the necessary background
for ADCs and DACs.

{\flushleft \em Digital to Analog Converters.}
An $n$-bit DAC converts an $n$-bit digital value into a corresponding analog
voltage. A typical DAC uses an R-2R network as shown in Figure~\ref{fig:r2r}.
Each stage in the network comprises of a pair of 
resistors (R and 2R) and a switch $s_i$ 
($0 \le i \le n$). The switch $s_i$ is closed only if the $i$-th bit of the
digital data is one. The resulting voltage at the amplifier's input is the
sum of voltages due to each closed switch. 
Each switch contributes a voltage of $\frac{1}{2^{n-i}}$, 
where $n < i \le 0$. 
Thus, the analog voltage at the amplifier's input
is proportional to the digital data.

\begin{figure}
\includegraphics[width=9cm]{R-2R-DAC.png}
\centering
\caption{Block diagram of an n-bit R-2R DAC\label{fig:r2r}}
\end{figure}

{\flushleft \em Analog to Digital Converters.}
An $n$-bit ADC converts an analog voltage (or current) to 
an equivalent $n$-bit digital data. The commonly used ADCs are
flash, successive-approximation, and sigma-delta ADCs. We give
a brief background on the successive-approximation ADC; the 
ADC we used for our experiments.

A successive-approximation ADC {\em SA-ADC} converts an analog
input to digital using a binary search technique.
In SA-ADC, a successive approximation register(SAR) containing 
a digital value is fed to an internal DAC and its output is 
compared to the input voltage. A sample and hold(SAH) circuit 
is used to supply constant input voltage to the comparator. 
A SA-ADC provides medium levels of accuracy and conversion time. 
The PUF output variation can occur because of microscopic variations 
in electrical components used in SAH circuit and internal DAC.
\begin{figure}[h]
\includegraphics[width=6cm]{SA-ADC.png}
\centering
\caption{Block diagram of an n-bit successive-approximation ADC}
\end{figure}
%It suffers from thermal noise for capacitors($kT/C$) which increases with temperature and is inversely proportional to the capacitance. However with higher capacitance, power consumption increases and accuracy decreases and usually a trade off is required.

\section{Mixed Signal PUFs}
This section describes how the MS-PUFs are built.

\subsection{Constructing the MS-PUF}
The MS-PUF is created by connecting outputs of DACs to 
input channels of an ADC. This creates loops between the processors
core, DACs, and ADCs as shown in Figure~\ref{fig:ms_puf}.
If there are $\alpha$ DACs and $\beta$ ADC channels in the device
(and $\alpha \le \beta$), then $\alpha$ such loops are made resulting 
in $\alpha$ feedback channels.
The PUFs challenge is a time-series pattern comprising of
$n$-bit digital values and a sequence of random channel identifiers.
The response is a time-series pattern comprising of $n$-bit digital 
values obtained from the ADC. The values for $n$ and $m$ are
respectively dependent on the DAC and ADC resolutions. Typically, $n = m$
for a device and have values 8, 12, 16, or 24.

Algorithm~\ref{algo:puf_sw} depicts the MS-PUF. 
The sofware running in the processor core 
selects the first channel from the sequence and 
transfers the time-series digital pattern to the corresponding
channel one value at a time ({\em line~5}). Each input is converted to an analog voltage (or
current) by the DAC and reconverted to a  digital value by the ADC.
The software program reads these values from the ADC to form a new digital time-series 
pattern ({\em line~7}). This pattern forms the input to the next channel in the sequence. The process
is repeated for the entire sequence. A function $g$ processes the initial pattern ($\mathbf C$) and 
the final pattern formed (corresponding to $\mathbf L_{len(L)}$) to form the PUFs response.

\begin{algorithm}[!ht] \label{algo:puf_sw}
\begin{scriptsize}
	\caption{\emph{MS-PUF : Obtaining the Response to a Challenge}}
	\DontPrintSemicolon
 	\LinesNumbered
	\KwIn{$(\mathbf C,\mathbf L)$ : (an array representing digital time-series pattern, 
                                         a random sequence of channel identifiers)}
	\KwOut{$\mathbf R$ : an array representing the response}
	\Begin{
		$\mathbf T \leftarrow \mathbf C$ \;
		\For{$i \in \{1, 2, \cdots, len(\mathbf L)\}$}{
			\For{$j \in \{1, 2, \cdots, len(\mathbf T)\}$}{
				Output\_to\_DAC($\mathbf T_j$, $\mathbf L_i$); \\
				/* Wait for ADC to finish conversion */ \;
				$r$ = Input\_from\_ADC($\mathbf L_i$); \;	
				$\mathbf T_j \leftarrow r$ \;
			}
		}
		$\mathbf R \leftarrow g(\mathbf C, \mathbf T)$ \;
		\Return ${\mathbf R}$
	}
\end{scriptsize}
\end{algorithm}

{\flushleft \em Variations in the MS-PUF Response.}
In every iteration of Algorithm~\ref{algo:puf_sw}, $r$, the $m$-bit 
reply from the ADC is the result of the $n$-bit DAC input $\mathbf T_j$ 
undergoing a conversion to analog followed by a reconversion to digital. 
Due to reasons such as quatification and ambient noise, the result ($r$)
would be different from $\mathbf T_j$ even if $m = n$.

Due to nano-scale variations, devices would inject certain randomness into
$r$ that would depend on the channel used and would vary from one device to another. 
There are several potential reasons for variations between channels and devices.
A DAC output for instance, is influenced by several nano-scale features such 
as voltage references, resistor tolerances, 
capacitive effects, and amplifier differences. This could result 
in two DACs having different responses to the same input. 
Successive approximation ADCs have built in DACs, therefore
would be affected by the same variations as the DAC. Additionally,
other components in the ADC such as sample and hold circuits and
comparators could lead to differences in the reply.

{\flushleft \em Preventing Modeling Attacks.}

\subsection{A Formal Model for MS-PUFs}

\subsection{xxx}

A R-2R DAC suffers from thermal noise which increases with resistance and temperature. The PUF output variation can occur because of microscopic variations in the resistors and switches used.

\section{Evaluation}

\section{Related Works}
In the last decade various types of PUFs have been proposed. The PUFs are categorized into either
electronic PUFs or non-electronic PUFs. Electronic PUFs are further classified into analog-electronic
PUFs and digital-intrinsic PUFs. The various types of PUFs are discussed in this section.

\subsection{Non-electronic PUFs}
As the name suggests, non-electronic PUFs use randomness arising due to non electronic components in the system
such as from optical, acoustic, and magnetic media. In the case of \textit{Optical PUFs} for instance, the main 
component is an optical token which is doped with refractive glass spheres in a random fashion. When the token is 
radiated with a laser, a speckle pattern is formed. The patterns are  unique because  of the microscopic differences 
between the tokens. In \textit{Acoustic PUFs}, mechanical vibrations is the main source of
randomness. The vibrations causes a sound wave, which scatters yielding unique reflections. 
Non-electronic PUFs do not have wide usability due to expensive hardware and accurate positioning 
required for their construction.

\subsection{Analog electronic PUFs}
In these PUFs, partial randomness introduced due to electronic quantities like capacitance or 
resistance is measured. 
For example, in a \textit{power distribution PUF}, voltage drops based on resistance variations
in the power grid of a chip is measured. The voltage drops are partial random owing to the
variations in manufacturing. In a \textit{V\_T PUFs}, differences due to manufacturing variations
on the threshold voltage of transistors is exploited. In both these PUFs external measuring instruments
are required, making them expensive.

\subsection{Digital Intrinsic PUFs}
The PUFs discussed so far are non-intrinsic, where randomness is explicitly introduced into the
system. In intrinsic PUFs, the challenge response 
collection mechanism is embedded in the device. There are two types of intrinsic PUFs :
delay-based and memory-based. In delay-based PUFs, random delays in propagating a current signal
through wires is exploited. Commonly used delay-based PUFs are arbiter PUFs and ring oscillator 
PUFs. In memory-based PUFs, randomness in settling of values in memory cells is exploited. 
These include SRAM PUFs, Butterfly PUFs and Flip-flop PUFs.

\section{Conclusion}